A fully labelled proof system for intuitionistic modal logics

### 2019-09-26 caleidoscope/caleidoscope.tex
cut elim

@@

>>>>

%\newcommand{\height}[1]{\mathrm{ht}(#1)}
\newcommand{\height}[1]{|#1|}

@@

\lutz{some start below}

In this section we are going to prove the admissibility of cut for $\labIKp$.

\begin{theorem}
  \label{thm:cut-adm}
  The rule $\labrn{cut}$ is admissible for $\labIKp$.
\end{theorem}

This theorem directly entails the implication \ref{ii}$\implies$\ref{iii} of Theorem~\ref{thm:cutfree-compl}. But before we can prove it, we need a series of auxiliary  lemmas.

The \emph{height} of a derivation $\DD$, denoted by $\height\DD$, is \todo{finish}

We say that a rule is \emph{height-preserving admissible} if for every derivation $\DD$ of its premise there is a derivation $\DD'$ of its conclusion such that $\height{\DD'}\le\height\DD$. A rule is \emph{height-preserving invertible} if \todo{...}

The following two lemmas are standard.

\begin{lemma}
  \label{lem:weak-adm}
  The weakening rule
  ~$\vlinf{\rn{weak}}{}{\lseq{\B,\B'}{\Left,\Left'}{\Right,\Right'}}{\lseq{\B}{\Left}{\Right}}$~
  is height-preserving admissible for $\labIKp$.
\end{lemma}

\begin{proof}
  \todo{}
\end{proof}

\begin{lemma}
  \label{lem:inv}
  The rules ..... are height-preserving invertible in
  $\labIKp$. Furthermore, the rule $\llabrn{\IMP}$ is
  height-preserving invertible for the right premise.
\end{lemma}

\begin{proof}
  \todo{}
\end{proof}

The next lemma looks like a special case of
Proposition~\ref{prop:mon-adm}, but it is not. First, we need to
preserve the height, and second, we cannot prove it using cut as we
need it to eliminate the cut.

\begin{lemma}
  The atomic version of $\llabrn{mon}$
  $$
  \vlinf{\labrn{mon_a}}{}{\lseq{\B, x \le x'}{\Left, \labels{x}{a}}\Right}{
    \lseq{\B, x \le x'}{\Left, \labels{x}{a}, \labels{x'}{a}}\Right}
  $$
  is height-preserving admissible for $\labIKp$.
\end{lemma}

\begin{proof}
  \todo{}
\end{proof}

The next lemma is the core of our cut elimination proof.

\begin{lemma}
  \label{lem:reduction}
  Given a derivation of shape
  $$
  \vlderivation{
    \vliiin{\labrn{cut}}{}{\lseq\B\Left\Right}{
      \vlhtr{\DD_1}{\lseq\B\Left{\Right,\labels{z}{C}}}}{
      \vlhy{}}{
      \vlhtr{\DD_2}{\lseq\B{\Left,\labels{z}{C}}{\Right}}}}
  $$
  where $\DD_1$ and $\DD_2$ are both cut-free, there is a cut-free
  derivation of ${\lseq\B\Left\Right}$
\end{lemma}

$\vlderivation{\vlhtr{\DD}{\lseq\B\Left\Right}}$
